The Rockerverse: Images, Packages, and Applications for Containerization with R

### 2020-01-23 manuscript.Rmd

    affiliation: University of Gothenburg, Sweden

@@

      - Department of Biological and Environmental Sciences

@@

      - 413 19 Gothenburg, Sweden

@@

**\pkg{outsider}** ([https://antonellilab.github.io/outsider/](https://antonellilab.github.io/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible.
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own R-based workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- informative examples of how modules work, just as a reminder for @nuest
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->

>>>>

    affiliation: Gothenburg Global Biodiversity Centre, Sweden

@@

@@

      - 413 19 Göteborg, Sweden

@@

**\pkg{outsider}** ([https://docs.ropensci.org/outsider/](https://docs.ropensci.org/outsider/)) tackles the problem of integrating external programs into an R workflow \citep{bennett_outsider_2020}.
Installation and usage of external programs can be difficult, convoluted and even impossible if the platform is incompatible.
Therefore \pkg{outsider} uses the platform-independent Docker images to encapsulate processes in _outsider modules_.
Each outsider module has a `Dockerfile` and an R package with functions for interacting with the encapsulated tool.
Using only R functions, an end-user can install a module with the \pkg{outsider} package and then call module code to integrate a tool into their own R-based workflow seamlessly.
The \pkg{outsider} package and module manage the containers and handle the transmission of arguments and the transfer of files to and from a container.
These functionalities also allow a user to launch module code on a remote machine via SSH, expanding the potential computational scale.
Outsider modules can be hosted code-sharing services, e.g., on GitHub, and \pkg{outsider} contains discovery functions for them.
<!-- informative examples of how modules work, just as a reminder for @nuest
https://github.com/DomBennett/om..astral/blob/master/examples/astral.R
https://github.com/DomBennett/om..raxml#partitioned-dna-analysis
-->
